Quantifying the rock-paper-scissors effect

### 2022-06-27 inst/rock_paper_scissors_monster.Rmd
hyper3 analysis

@@

```{r,label=maxpHWM,cache=TRUE}
maxp(HWM)
```

@@

```{r,label=specicifgtest,cache=TRUE}
specificp.gt.test(HWM,"M",0)
```

@@

>>>>

(takes about an hour to run without cache)

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

@@

```{r,label=maxpHWM,cache=TRUE}
maxHW <- maxp(HWM)
maxHW
```

@@

```{r,label=specificgtesthwm,cache=TRUE}
specificp.gt.test(HWM,"M",0)
```

@@

### Analysis of seven-tooth gear with `hyper3`

We can now use `hyper3` to analyse the same dataset.  Here the
likelihood function for pairwise competition between $p_1$ and $p_2$,
with $a$ procyclic wins and $b$ anticyclic wins, is

$$
\left(\frac{\lambda p_1}{\lambda p_1 + p_2}\right)^a
\cdot
\left(\frac{p_2}{\lambda p_1 + p_2}\right)^b
$$

where $\lambda$ is a quantification of the asymmetry.  We can modify
the `hyper2` idiom to make a `hyper3` object instead:

```{r,label=defineh3seven}
H3seven <- function(lambda){
  H3 <- hyper3()
  for(i in 1:6){
     jj <- 1  # diagonal players do not have lambda-advantage
     names(jj) <- tt[i]
     H3[jj] %<>% inc(hw[i,i])

     jj <- lambda
     names(jj) <- tt[i+1]
     H3[jj] %<>% inc(hw[i,i+1])

     jj <- c(1,lambda)
     names(jj) <- c(tt[i],tt[i+1])
     H3[jj] %<>% dec(hw[i,i] + hw[i,i+1])
  }  # i loop closes

  ## Now do final one:
  jj <- 1
  names(jj) <- tt[7]
  H3[jj] %<>% inc(hw[7,7])

  jj <- lambda
  names(jj) <- tt[1]
  H3[jj] %<>% inc(hw[7,1])

  jj <- c(1,lambda)
  names(jj) <- c(tt[7],tt[1])
  H3[jj] %<>% dec(hw[7,7] + hw[7,1])

  return(H3)
}  # function H3seven() defn closes
```

Now we can execute it:

```{r,label=executeh3seven}
H3 <- H3seven(lambda=1.66)
H3
```

And try a profile likelihood:

```{r,label=defineproflikeh3}
profl <- function(lambda){
  H <- H3seven(lambda)
  return(loglik(maxp(H),H))
}
```

```{r,label=proflikh3,cache=TRUE}
lam <- seq(from=0.8,to=2.6,len=15)
like3 <- sapply(lam,profl)
like3 <- like3-max(like3)
```

```{r,plotproflike3}
plot(lam,like3,type='b')
abline(h=c(0,-2))
abline(v=1)
```
